A semantics of core AUTOSAR

### 2016-04-24 paper/paper.tex
Some TODOs.
More work on the eprover to LaTeX conversion
Preparing macro merge by moving them to a separate file
LaTeX macros merged from paper and semantics. Next step i formatting.

\documentclass[twocolumn]{article}
\usepackage{verbatim}
\usepackage[utf8x]{inputenc}
\usepackage[a4paper,margin=1.4cm]{geometry}
\usepackage[fleqn]{amsmath}
\usepackage{mathpartir}
% \usepackage{minted}
\title{A semantics of core AUTOSAR}
\author{Johan Nordlander \and Patrik Jansson}

\newcommand{\V}[1]{\text{\it #1}}
\newcommand{\Code} {\V{Code}}
\newcommand{\Cont} {\V{Cont}}
\newcommand{\State}{\V{State}}
\newcommand{\Act}  {\V{Act}}
\newcommand{\Upd}  {\V{Upd}}
\newcommand{\Xs}   {\V{Xs}}
\newcommand{\Vs}   {\V{Vs}}
\newcommand{\Cs}   {\V{Cs}}

\newcommand{\kw}[1]{\textbf{#1}}
\newcommand{\void}    {\kw{void}}
\newcommand{\nil}     {\kw{nil}}
\newcommand{\true}    {\kw{true}}
\newcommand{\false}   {\kw{false}}
\newcommand{\idle}    {\kw{idle}}
\newcommand{\pending} {\kw{pending}}
\newcommand{\serving}[2]{\kw{serving}(#1,#2)}

\newcommand{\adr}[1]{\emph{#1}}
\newcommand{\sep}{\hspace{2ex}|\hspace{2ex}}

\newcommand{\runnable}[4]{\kw{runnable}(#1,#2,#3,#4)}
\newcommand{\rinst}   [4]{\kw{rinst}(#1,#2,#3,#4)}
\newcommand{\excl}    [2]{\kw{excl}(#1,#2)}
\newcommand{\irv}     [2]{\kw{irv}(#1,#2)}
\newcommand{\qelem}   [2]{\kw{qelem}(#1,#2)}
\newcommand{\delem}   [3]{\kw{delem}(#1,#2,#3)}
\newcommand{\opres}   [2]{\kw{opres}(#1,#2)}
\newcommand{\timer}   [3]{\kw{timer}(#1,#2, #3)}

\newcommand{\rte}[1]{\text{rte\_#1}}
\newcommand{\Enter}     [2]{\rte{Enter}(#1,#2)}
\newcommand{\Exit}      [2]{\rte{Exit}(#1,#2)}
\newcommand{\IrvWrite}  [3]{\rte{IrvWrite}(#1,#2,#3)}
\newcommand{\IrvRead}   [2]{\rte{IrvRead}(#1,#2)}
\newcommand{\Send}      [4]{\rte{Send}(#1,#2,#3,#4)}
\newcommand{\Receive}   [3]{\rte{Receive}(#1,#2,#3)}
\newcommand{\Write}     [4]{\rte{Write}(#1,#2,#3,#4)}
\newcommand{\Read}      [3]{\rte{Read}(#1,#2,#3)}
\newcommand{\Invalidate}[3]{\rte{Invalidate}(#1,#2,#3)}
\newcommand{\IsUpdated} [3]{\rte{IsUpdated}(#1,#2,#3)}
\newcommand{\Call}      [4]{\rte{Call}(#1,#2,#3,#4)}
\newcommand{\Result}    [3]{\rte{Result}(#1,#2,#3)}
\newcommand{\Return}    [1]{\text{return}(#1)}

\newcommand{\red} [1]{\xrightarrow{#1}}
\newcommand{\say} [2]{#1 ! #2}
\newcommand{\hear}[2]{#1 ? #2}

\newcommand{\ent}{\text{\small ent}}
\newcommand{\ex} {\text{\small ex}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt plus 2pt minus 1pt}

@@

The process terms of our calculus are as follows:
\begin{itemize}
\item $\runnable{\adr{R:I}}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\rinst{\adr{R:I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\excl{\adr{X:I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\irv{\adr{S:I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\delem{\adr{E:U:I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\qelem{\adr{E:U:I}}{\Vs}$ \newline
Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
\item $\opres{\adr{O:U:I}}{\Vs}$ \newline
Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
\item $\timer{\adr{R:I}}{T_p}{T}$ \newline
Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
\end{itemize}

@@

The code part of a runnable instance should technically be a representation of the C or Matlab function that must accompany an AUTOSAR runnable definition once it is complete. However, since our task in this report is not to investigate the semantics of these languages in every detail, a more abstract notion of executable code will be required. To this end, we have chosen to ignore all details of internal computations, and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls. Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
$$
\begin{array}{lll}
	\Code	& ::=	& \Enter{X}{\Cont}			\\
			& |	& \Exit{X}{\Cont}			\\
			& |	& \IrvWrite{S}{V}{\Cont}		\\
			& |	& \IrvRead{S}{\Cont}			\\
			& |	& \Send{E}{U}{V}{\Cont}		\\
			& |	& \Receive{E}{U}{\Cont}		\\
			& |	& \Write{E}{U}{V}{\Cont}		\\
			& |	& \Read{E}{U}{\Cont}		\\
			& |	& \Invalidate{E}{U}{\Cont}		\\
			& |	& \IsUpdated{E}{U}{\Cont}		\\
			& |	& \Call{O}{U}{V}{\Cont}		\\
			& |	& \Result{O}{U}{\Cont}		\\
			& |	& \Return{V}				\\
	\\
	\Cont	& \in	& V \rightarrow \Code		\\
\end{array}
$$

Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
$$
\begin{array}{rcll}
	P,Q 	& ::= &
		\runnable{A}{T}{\Act}{N}	\\
		& | &
		\rinst{A}{C}{\Xs}{\Code}	\\
		& | &
		\excl{A}{B}			\\
		& | &
		\irv{A}{V}				\\
		& | &
		\delem{A}{B}{V}			\\
		& | &
		\qelem{A}{\Vs}			\\
		& | &
		\opres{A}{\Vs}			\\
		& | &
		\timer{A}{T_p}{T}		\\
		& | &
		P|Q					\\
\end{array}
$$

@@

\item For each runnable $R$ of each component $I$, a term
$$
\begin{array}{c}
	\runnable{\adr{R:I}}{0.0}{\Act}{0}
\end{array}
$$
where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\idle$.
% SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable

@@

\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\qelem{\adr{E:U:I}}{\epsilon}
\end{array}
$$
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
$$
\begin{array}{c}
	\delem{\adr{E:U:I}}{\false}{\text{NO\_DATA}}
\end{array}
$$

@@

@@

\input{semantics.tex}

>>>>

\documentclass[twocolumn]{article}
\usepackage{verbatim}
\usepackage[utf8x]{inputenc}
\usepackage[a4paper,margin=1.4cm]{geometry}
\usepackage[fleqn]{amsmath}
\usepackage{mathpartir}
% \usepackage{minted}
\title{A semantics of core AUTOSAR}
\author{Johan Nordlander \and Patrik Jansson}
%%%%%%%
\input{macros.tex}
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt plus 2pt minus 1pt}

@@

The process terms of our calculus are as follows:
\begin{itemize}
\item $\runnable{\adr{R:I}}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\rinst{\adr{R:I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\excl{\adr{X:I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\irv{\adr{S:I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\delem{\adr{E:U:I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\qelem{\adr{E:U:I}}{\Vs}{TODO}$ \newline
Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
\item $\opres{\adr{O:U:I}}{\Vs}$ \newline
Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
\item $\timer{\adr{R:I}}{T_p}{T}$ \newline  %TODO: timer has only 2 args in the semantics
Timer state for runnable $R$ of component $I$ and period $T_p$, with $T$ seconds left of its current cycle.
\end{itemize}

@@

The code part of a runnable instance should technically be a representation of the C or Matlab function that must accompany an AUTOSAR runnable definition once it is complete. However, since our task in this report is not to investigate the semantics of these languages in every detail, a more abstract notion of executable code will be required. To this end, we have chosen to ignore all details of internal computations, and just capture the observable effects of runnable execution as a pure sequence of RTE (VFB) calls. Because the result from an RTE call may in general affect subsequent runnable behavior, we use a continuation-passing style to allow any data-dependencies to be expressed.
%TODO: several LaTeX commands here are used with more arguments than in the semantics below
$$
\begin{array}{lll}
	\Code	& ::=	& \Enter{X}{\Cont}			\\
			& |	& \Exit{X}{\Cont}			\\
			& |	& \IrvWrite{S}{V}{\Cont}		\\
			& |	& \IrvRead{S}{\Cont}			\\
			& |	& \Send{E}{U}{V}{\Cont}		\\
			& |	& \Receive{E}{U}{\Cont}		\\
			& |	& \Write{E}{U}{V}{\Cont}		\\
			& |	& \Read{E}{U}{\Cont}		\\
			& |	& \Invalidate{E}{U}{\Cont}		\\
			& |	& \IsUpdated{E}{U}{\Cont}		\\
			& |	& \Call{O}{U}{V}{\Cont}		\\
			& |	& \Result{O}{U}{\Cont}		\\
			& |	& \Return{V}				\\
	\\
	\Cont	& \in	& V \rightarrow \Code		\\
\end{array}
$$

Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
$$
\begin{array}{rcll}
	P,Q 	& ::= &
		\runnable{A}{K}{T}{\Act}{N}	\\
		& | &
		\rinst{A}{C}{\Xs}{\Code}	\\
		& | &
		\excl{A}{B}			\\
		& | &
		\irv{A}{V}				\\
		& | &
		\delem{A}{B}{V}			\\
		& | &
		\qelem{A}{\Vs}{TODO}			\\
		& | &
		\opres{A}{\Vs}			\\
		& | &
		\timer{A}{T_p}{T}		\\
		& | &
		P|Q					\\
\end{array}
$$

@@

\item For each runnable $R$ of each component $I$, a term
$$
\begin{array}{c}
	\runnable{\adr{R:I}}{TODO}{0.0}{\Act}{0}
\end{array}
$$
where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\pending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\idle$.
% SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable

@@

\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\qelem{\adr{E:U:I}}{\epsilon}{TODO}
\end{array}
$$
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
$$
\begin{array}{c}
	\delem{\adr{E:U:I}}{\false}{\text{NO\_DATA}}
\end{array}
$$

@@

Rule and examples of provided delegation port compatibility don't match

Multiple inconsistent CompSpecs

@@

\input{../prolog/tool/semantics.tex}
